### 2018-04-21 Manuscript/chapter4.tex
Final draft for thesis

Somewhat surprisingly, the transcriptomic response to virus infection in our experiment was fairly limited. We found only 43 transcripts to be differentially expressed, some with known immune functions (Table \ref{tbl:virusGenes}) such as argonaute-2 and a gene with similarity to MD-2 lipid recognition protein, as well as additional genes related to transcriptional regulation, and muscle contraction. The small number of DEGs in this study may be partly explained by the large amount of noise in the data (Figures \ref{fig:mdsPlots}B, \ref{fig:pcpRutterVirus}, \ref{fig:litreClusterRutter}, \ref{fig:RutterSM1}, \ref{fig:RutterSM2}, \ref{fig:RutterSM3}, \ref{fig:RutterSM4}).

Given the noisy nature of our data, and our desire to hone in on genes with real expression differences, we compared our data to the Galbraith study (\citealt{galbraith}), which also examined bees response to viral infection. In contrast to our study, Galbraith et al. identified a large number of virus responsive trancripts, and generally had less noise in their data (Figures \ref{fig:mdsPlots}A, \ref{fig:pcpGalbraith}, \ref{fig:litreCluster2}, \ref{fig:litreCluster1}, \ref{fig:GalbraithClust1SM}, \ref{fig:GalbraithClust2SM}, \ref{fig:GalbraithClust3SM}, \ref{fig:GalbraithClust4SM}).  To identify the most reliable virus-responsive genes from our study, we looked for overlap in the DEGs associated with virus infection on both experiments. We found a large, statistically significant (p-value < 2.2e-16) overlap, with 26/38 (68\%) of virus-responsive DEGs from our study also showing response to virus infection in Galbraith et al. (Figure \ref{fig:GRVenn}). This result gives us confidence that, although noisy, we were able to uncover consistent, replicable gene expression responses to virus infection with our data.

** paragraph here to talk about the importance and utility of data visualization in RNA-seq data.  For Lindsay to write.

@@

Overall, these data suggest complex transcriptomic responses to multiple stressors in honey bees. Diet has the potential for large and profound effects on transcriptional responses in honey bees, and differences in diet may set up the potential for both resistance and resilience to virus infection.

\section{Appendix}

>>>>

Somewhat surprisingly, the transcriptomic response to virus infection in our experiment was fairly limited. We found only 43 transcripts to be differentially expressed, some with known immune functions (Table \ref{tbl:virusGenes}) such as argonaute-2 and a gene with similarity to MD-2 lipid recognition protein, as well as additional genes related to transcriptional regulation, and muscle contraction. The small number of DEGs in this study may be partly explained by the large amount of noise in the data (Figures \ref{fig:mdsPlots}B, \ref{fig:pcpRutterVirus}, \ref{fig:litreClusterRutter}, \ref{fig:RutterSM1}, \ref{fig:RutterSM2}, \ref{fig:RutterSM3}, and \ref{fig:RutterSM4}).

Given the noisy nature of our data, and our desire to hone in on genes with real expression differences, we compared our data to the Galbraith study (\citealt{galbraith}), which also examined bees response to viral infection. In contrast to our study, Galbraith et al. identified a large number of virus responsive trancripts, and generally had less noise in their data (Figures \ref{fig:mdsPlots}A, \ref{fig:pcpGalbraith}, \ref{fig:litreCluster2}, \ref{fig:litreCluster1}, \ref{fig:GalbraithClust1SM}, \ref{fig:GalbraithClust2SM}, \ref{fig:GalbraithClust3SM}, and \ref{fig:GalbraithClust4SM}). To identify the most reliable virus-responsive genes from our study, we looked for overlap in the DEGs associated with virus infection on both experiments. We found a large, statistically significant (p-value < 2.2e-16) overlap, with 26/38 (68\%) of virus-responsive DEGs from our study also showing response to virus infection in Galbraith et al. (Figure \ref{fig:GRVenn}). This result gives us confidence that, although noisy, we were able to uncover consistent, replicable gene expression responses to virus infection with our data.

The DESeq2 package comes with certain visualization options that are popular in RNA-sequencing analysis. One of these visualization is the multidimensional scaling (MDS) plot, which allows users to visualize the similarity between samples within a dataset. We could determine from this plot that indeed the Galbraith data may show more similarity between its replicates and differences between its treatments compared to our data (Figure \ref{fig:mdsPlots}). However, the MDS plot only shows us information at the sample level. We wanted to investigate how these differences in the signal:to:noise ratios of the datasets would affect the structure of any resulting DEGs. As a result, we also used three plotting techniques from the bigPint package: We investigated the 1,019 virus-related DEGs from the Galbraith dataset and the 43 virus-related DEGs from our dataset using parallel coordinate lines, litre plots, and scatterplot matrices. To prevent overlapping issues in our plots, we used a hierarchical clustering technique for the parallel coordinate lines to separate the set of DEGs into smaller groups. We also needed to examine four subsets of samples from the Galbraith dataset to make effective use of the scatterplot matrices. After these tailorizations, we determined that the same patterns we saw in the MDS plots regarding the entire dataset extended down the pipeline analysis into the DEG calls: Even the DEGs from the Galbraith datasaet showed more similarity between their replicates and differences between their treatments compared to those from our data. However, the 365 DEGs from the Galbraith data in Cluster 2 of Figure \ref{fig:pcpGalbraith} showed an inconsistent first replicate in the treatment group (``T.1''), which was something we observed in the MDS plot. This indicates that this feature also extended down the analysis pipeline into DEG calls.

@@

Overall, these data suggest complex transcriptomic responses to multiple stressors in honey bees. Diet has the potential for large and profound effects on transcriptional responses in honey bees, and differences in diet may set up the potential for both resistance and resilience to virus infection. Moreover, this study in general also demonstrated the possible benefits of examining multiple datasets to address inherently messy biological data. For instance, by verifying the subsantial overlap in our DEG lists to those obtained in another study that addressed a similar question but in a more controlled manner, we were able to feel more confident in our DEG calls.
